Causal Discovery on Precarity and Depression

### 2024-12-28 draft_v1/draft_v1.qmd
add_presum_graph

```{r, include=FALSE}
# Load the dataset 
dat <- read_sav("../data/HELIUS_LEONIE.sav")

@@

This setup yields 16 combinations (2 significance levels × 4 thresholds × 2 CI tests), applied across three algorithms, with each combination repeated for 100 bootstrap samples, resulting in a total of 1,600 resulting graphs. For analyses involving individual symptom variables, we streamline the setup by using thresholds of 0.6 and 0.7, reducing the number of bootstrap samples to 30. Additionally, we fix the skeleton of edges among symptom variables based on the common structure estimated across all three algorithms, thereby reducing the computational time required to estimate skeletons. This fixed structure aligns with commonly observed skeleton structures in the literature, ensuring relevance to existing findings (**cite the comp_model paper**).[^2]

To summarize the results, we identify the most frequently occurring edge endpoints across different experimental setups. This ensures that only stable and consistent edges are retained, providing a clearer and more reliable understanding of the relationships between precarity factors and depression. Refer to @fig-workflow for the analysis workflow.

[^2]: 
To reduce spurious edges in the symptom network, we fixed the skeleton of symptom interactions using a common structure derived from PC, FCI, and CCI algorithms with an alpha level of 0.001. This approach maintained sparsity in symptom interactions while allowing the free estimation of causal directions between symptoms and precarity factors, balancing computational efficiency with meaningful results.

@@

<!-- Centrality and Connectivity -->
Certain nodes in the network emerge as more *causally* central, highlighting their potential importance as intervention points. Among the depressive symptoms, *dep*, *slp*, and *glt* emerge as particularly influential, with multiple outgoing edges (o->) to other symptoms,  suggesting their roles as potential key drivers within the symptom network. Especially, symptoms such as *glt*, *slp*, and *anh* are particularly critical, potentially acting as initiator or activator nodes within the network due to their apparent connections with precarity factors. These symptoms appear to be especially sensitive to external stressors, potentially manifesting early in response to such conditions and subsequently activating other interconnected symptom nodes.
<!-- This sensitivity indicates their potential as early warning signs or key targets for interventions aimed at preventing the progression to severe depression. -->

@@

## Summarized Stable Edges Proportion

>>>>

```{r, include=FALSE}
# Load the dataset 
dat <- read_sav("../../data/HELIUS_LEONIE.sav")

@@

This setup yields 16 combinations (2 significance levels × 4 thresholds × 2 CI tests), applied across three algorithms, with each combination repeated for 100 bootstrap samples, resulting in a total of 1,600 resulting graphs. For analyses involving individual symptom variables, we streamline the setup by using thresholds of 0.6 and 0.7, reducing the number of bootstrap samples to 30. 
<!-- Additionally, we fix the skeleton of edges among symptom variables based on the common structure estimated across all three algorithms, thereby reducing the computational time required to estimate skeletons. This fixed structure aligns with commonly observed skeleton structures in the literature, ensuring relevance to existing findings (**cite the comp_model paper**).[^2]  -->

To summarize the results, we identify the most frequently occurring edge endpoints across different experimental setups.[^3] This ensures that only stable and consistent edges are retained, providing a clearer and more reliable understanding of the relationships between precarity factors and depression. Refer to @fig-workflow for the analysis workflow.

<!-- [^2]:  -->
<!-- To reduce spurious edges in the symptom network, we fixed the skeleton of symptom interactions using a common structure derived from PC, FCI, and CCI algorithms with an alpha level of 0.001. This approach maintained sparsity in symptom interactions while allowing the free estimation of causal directions between symptoms and precarity factors, balancing computational efficiency with meaningful results. -->

[^3]:
The detailed proportion of each edge endpoint occurrence is shown in [@sec-propmatrix].

@@

<!-- Centrality and Connectivity -->
Certain nodes in the network emerge as more *causally* central, highlighting their potential importance as intervention points. Among the depressive symptoms, *anh*, *dep*, *slp*, and *glt* emerge as particularly influential, with multiple outgoing edges (o->) to other symptoms,  suggesting their roles as potential key drivers within the symptom network. Especially, symptoms such as *glt*, *slp*, and *anh* are particularly critical, potentially acting as initiator or activator nodes within the network due to their apparent connections with precarity factors. These symptoms appear to be especially sensitive to external stressors, potentially manifesting early in response to such conditions and subsequently activating other interconnected symptom nodes.
<!-- This sensitivity indicates their potential as early warning signs or key targets for interventions aimed at preventing the progression to severe depression. -->

@@

## Summarized Stable Edges Proportion {#sec-propmatrix}

### 2024-12-28 draft_v1/draft_v1.tex
add_presum_graph

This setup yields 16 combinations (2 significance levels × 4 thresholds
× 2 CI tests), applied across three algorithms, with each combination
repeated for 100 bootstrap samples, resulting in a total of 1,600
resulting graphs. For analyses involving individual symptom variables,
we streamline the setup by using thresholds of 0.6 and 0.7, reducing the
number of bootstrap samples to 30. Additionally, we fix the skeleton of
edges among symptom variables based on the common structure estimated
across all three algorithms, thereby reducing the computational time
required to estimate skeletons. This fixed structure aligns with
commonly observed skeleton structures in the literature, ensuring
relevance to existing findings (\textbf{cite the comp\_model
paper}).\footnote{To reduce spurious edges in the symptom network, we
  fixed the skeleton of symptom interactions using a common structure
  derived from PC, FCI, and CCI algorithms with an alpha level of 0.001.
  This approach maintained sparsity in symptom interactions while
  allowing the free estimation of causal directions between symptoms and
  precarity factors, balancing computational efficiency with meaningful
  results.}

To summarize the results, we identify the most frequently occurring edge
endpoints across different experimental setups. This ensures that only
stable and consistent edges are retained, providing a clearer and more
reliable understanding of the relationships between precarity factors
and depression. Refer to Figure~\ref{fig-workflow} for the analysis
workflow.

@@

Certain nodes in the network emerge as more \emph{causally} central,
highlighting their potential importance as intervention points. Among
the depressive symptoms, \emph{dep}, \emph{slp}, and \emph{glt} emerge
as particularly influential, with multiple outgoing edges
(o-\textgreater) to other symptoms, suggesting their roles as potential
key drivers within the symptom network. Especially, symptoms such as
\emph{glt}, \emph{slp}, and \emph{anh} are particularly critical,
potentially acting as initiator or activator nodes within the network
due to their apparent connections with precarity factors. These symptoms
appear to be especially sensitive to external stressors, potentially
manifesting early in response to such conditions and subsequently
activating other interconnected symptom nodes.

@@

\subsection{Summarized Stable Edges
Proportion}\label{summarized-stable-edges-proportion}

>>>>

This setup yields 16 combinations (2 significance levels × 4 thresholds
× 2 CI tests), applied across three algorithms, with each combination
repeated for 100 bootstrap samples, resulting in a total of 1,600
resulting graphs. For analyses involving individual symptom variables,
we streamline the setup by using thresholds of 0.6 and 0.7, reducing the
number of bootstrap samples to 30.

To summarize the results, we identify the most frequently occurring edge
endpoints across different experimental setups.\footnote{The detailed
  proportion of each edge endpoint occurrence is shown in
  Section~\ref{sec-propmatrix}.} This ensures that only stable and
consistent edges are retained, providing a clearer and more reliable
understanding of the relationships between precarity factors and
depression. Refer to Figure~\ref{fig-workflow} for the analysis
workflow.

@@

Certain nodes in the network emerge as more \emph{causally} central,
highlighting their potential importance as intervention points. Among
the depressive symptoms, \emph{anh}, \emph{dep}, \emph{slp}, and
\emph{glt} emerge as particularly influential, with multiple outgoing
edges (o-\textgreater) to other symptoms, suggesting their roles as
potential key drivers within the symptom network. Especially, symptoms
such as \emph{glt}, \emph{slp}, and \emph{anh} are particularly
critical, potentially acting as initiator or activator nodes within the
network due to their apparent connections with precarity factors. These
symptoms appear to be especially sensitive to external stressors,
potentially manifesting early in response to such conditions and
subsequently activating other interconnected symptom nodes.

@@

\subsection{Summarized Stable Edges Proportion}\label{sec-propmatrix}
